---
title: "Coursera Regression - Week 2 Materials"
author: "Scott Smith"
date: "Wednesday, December 10, 2014"
output: html_document
---

Week 2 Lecture Materials

###Statistical Regression Model Examples
Diamond Example
```{r, echo=TRUE}
library(UsingR)
data(diamond)
plot(diamond$carat,diamond$price,
     xlab="Mass (carats)",
     ylab = "Price (SIN $)",
     bg = "lightblue",
     col = "black", cex = 1.1, pch = 21, frame = FALSE)
abline(lm(price ~ carat, data = diamond),lwd = 2)
```

Fitting the model
```{r, echo=TRUE}
library(UsingR); data(diamond)
fit <- lm(price ~ carat, data = diamond)
coef(fit)
```

Getting more interpretable intercept
```{r, echo = TRUE}
library(UsingR); data(diamond)
fit2 <- lm(price ~ I(carat - mean(carat)), data = diamond)
coef(fit2)
```

Changing the scale
```{r, echo=TRUE}
library(UsingR); data(diamond)
fit3 <- lm(price ~ I(carat * 10), data = diamond)
coef(fit3)
```

Predicting price of the diamond
```{r,echo=TRUE}
library(UsingR); data(diamond)
fit <- lm(price ~ carat, data = diamond)
newx <- c(0.16, 0.27, 0.34)
coef(fit)[1] + coef(fit)[2]*newx
predict(fit, newdata = data.frame(carat = newx))
```

Residuals
```{r}
library(UsingR); data(diamond)
y <- diamond$price; x<- diamond$carat; n <- length(y)
fit <- lm(y~x)
e <- resid(fit)
max(abs(e - (y - yhat)))
max(abs(e - (y - coef(fit)[1] - coef(fit)[2] * x)))
plot(x,e,
     xlab = "Mass (Carats)",
     ylab = "Residuals (SIN $)",
     cex=1)
abline(0,0)
```

Tests for Regression Coefficients
```{r}
library(UsingR); data(diamond)
y<- diamond$price; x <- diamond$carat; n <- length(y)
beta1 <- cor(y,x) * sd(y) / sd(x)
beta0 <- mean(y) - beta1 * mean(x)
e <- y - beta0 - beta1 * x
sigma <- sqrt(sum(e^2) / (n-2))
ssx <- sum((x - mean(x))^2)
seBeta0 <- (1 / n + mean(x) ^ 2 / ssx) ^ .5 * sigma
seBeta1 <- sigma / sqrt(ssx)
tBeta0 <- beta0 / seBeta0; tBeta1 <- beta1 / seBeta1
pBeta0 <- 2 * pt(abs(tBeta0), df = n - 2, lower.tail = FALSE)
pBeta1 <- 2 * pt(abs(tBeta1), df = n - 2, lower.tail = FALSE)
coefTable <- rbind(c(beta0, seBeta0, tBeta0, pBeta0), c(beta1, seBeta1, tBeta1, pBeta1))
colnames(coefTable) <- c("Estimate", "Std. Error", "t value", "P(>|t|)")
rownames(coefTable) <- c("(Intercept)", "x")
coefTable
print("Alternate Approach")
fit <- lm(y~x)
summary(fit)$coefficients
print("Getting Confidence Interval")
sumCoef <- summary(fit)$coefficients
sumCoef[1,1] + c(-1,1) * qt(.975, df = fit$df) * sumCoef[1,2]
sumCoef[2,1] + c(-1,1) * qt(.975, df = fit$df) * sumCoef[2,2]
print("With 95% confidence, estimate that a 0.1 carat increase in size results in a 355.6 to 388.6 increase in price (SIN $).")
```

Prediction Interval
```{r}
newdata <- data.frame(x = xVals)
p1 <- predict(fit, newdata, interval = ("confidence"))
p2 <- predict(fit, newdata, interval = ("prediction"))
plot(x, y, frame=FALSE, xlab = "Carat", ylab  = "Dollars", pch = 21, col = "black", bg = "lightblue", cex = 2)
abline(fit, lwd = 2)
lines(xVals, p1[,2]); lines(xVals, p1[,3])
lines(xVals, p2[,2]); lines(xVals, p2[,3])
```


###Quiz 2
Question 1 & 2
```{r}
x <- c(0.61, 0.93, 0.83, 0.35, 0.54, 0.16, 0.91, 0.62, 0.62)
y <- c(0.67, 0.84, 0.6, 0.18, 0.85, 0.47, 1.1, 0.65, 0.36)
fit <- lm(y~x); summary(fit)
```

Question 3
```{r}
library(UsingR); data(mtcars)
fit <- lm(mtcars$mpg, mtcars$wt)
sumCoef <- summary(fit)$coefficients